Monte Carlo Bootstrap: current developments.

### 2019-02-05 draft.tex
started adding chi restarter. Need to tune tolerance.

\section{General Intuition of the MC bootstrap}
One can consider \ref{bootstrapequation} as a function of the spectrum and the
OPE coefficients, which should only be satisfied for physical conformal data.
Since any numerical approach involves truncating the OPE up to a maximum
dimension $\Delta^*$, it is important to estimate the remainder of the series in
order to bound the amount by which \ref{bootstrapequation} might be violated in
a given approximation. This was done by Rychkov et al,
who found the following error estimate holds (for $\Delta_*\gg 1$):
\begin{equation}
\Big|\sum_{(\Delta\geq \Delta_*),l}p_{\mathcal{O}} \, g_{\Delta,l}(z,\bar z)\Big| \, \leq \, \mathcal{R}(z,\bar z)\,,
\label{finalest4}
\end{equation}
where
\begin{equation}
\mathcal{R}(z,\bar z) \,\equiv \, \frac{(-\log|\rho(z)|)^{-4\Delta_\phi+1} \, 2^{4\Delta_\phi+1}}{\Gamma(4\Delta_\phi)} \frac{\Gamma(4\Delta_\phi, - \Delta_* \log |\rho(z)|)}{{(1-|\rho(z)|^2)}}\,.
\label{finalest5a}
\end{equation}

@@

The straightforward path is to define a cost function
\[
  C\sim | \mathcal{M}\cdot \vec{\rho} - \vec{\sigma} |
\]
and minimize it stochastically.
This naive approach proved to be untractable in the early stages of the
research, mostly because it became exponentially time consuming to look for more
operators. Thus, we decided to split the optimization in two steps. One that
finds an approximate spectrum without worrying about the OPE coefficients, and
then solving for the complete conformal data with a more refined Monte Carlo.
The first step was motivated by Gliozzi's determinant bootstrap, which we review
in the following section.

@@

\section{LogDet implementation}
\subsection{naive approach}
Given a putative spectrum, the determinant of \ref{matrix-op}
should get closer to zero as we approach a physically sensible set of operator
dimensions. Since this quantity could range over many orders of magnitude, we
decided to square it and then take the $\log$ of it. This proved to be
numerically troubling, since the behaviour of the logarithm around $0$ is far
from differentiable and the landscape was thus full of spurious minima out of
which the Metropolis algorithm could not get out.
\subsection{Cross smearing}
In order to smooth this behaviour out, we implemented an empirical
renormalization that consists in computing the value of the determinant at
adjacent points to the main one. Thus
\[
\begin{aligned}
  S({\Delta_i,l_i})& = \log\Det^2\mathcal{M}(\{\Delta_i,l_i\})\\
  \to
  S_{\mathrm{cross}}(\{\Delta_i,l_i\})  & =   \log\Det^2\mathcal{M}(\{\Delta_i,l_i\}) + \sum_i\left(
  \log\Det^2\mathcal{M}(\{\Delta_i+\delta,l_i\})  +
  \log\Det^2\mathcal{M}(\{\Delta_i-\delta,l_i\})  \right)
\end{aligned}
\]

This action was much more efficient at finding the correct dimensions for the
free theory, and numerical exploration showed that the basin of attraction for
each given operator's dimension was very sharp, as long as the other ones were
kept at the exact value. Small perturbations destroyed this basin rather
quickly, and thus it was hard to get the MC routine to solve for a large number
of operators  ($\sim 7$)at the same time.

@@

The next breakthrough in the routine came when we decided to start with few
($\sim 4$) operators and add the next operators one by one. To avoid spoiling
the previous solution, only the new operator was varied for a given number of
steps before we allowed the whole spectrum to be optimized. This gave a very
efficient routine with a large radius of convergence.

@@

\section{$\chi^2$ refinement}
\subsection{cWLS as a way to determine OPEs}
Since the dependence of \ref{cross-check} on $\vec{\rho}$ is linear, for
$N_z>N_{Op}$ it is possible to find the set of OPE coefficients that best
satisfy crossing via the Weighted Least Squares method\footnote{Should we add a
reference here, other than Wikipedia?}. For a generic spectrum
obtained via the method described in section \ref{rec-check}, this naive set of OPE
coefficients will have some negative components, which stems from the
fact that our spectrum is only approximate. Since the error estimate depends
crucially on the positivity of the OPE coefficients, we need to find the
solution that minimizes \ref{cross-check} and has all entries non-negative. This
is attained by setting the first negative OPE coefficient (from back to forth,
if $\Delta_i<\Delta_{i+1}$) and then running a WLS routine with the remaining
entries. If this solution still has negative entries, the proccedure is repeated
until an acceptable $\vec{\rho}$ is obtained.

>>>>

\section{General Intuition of the MC bootstrap}
One can consider \ref{bootstrapequation} as a function of the spectrum and the
OPE coefficients, which should only be satisfied for physical conformal data.
Since almost any numerical approach involves truncating the OPE up to a maximum
dimension $\Delta^*$, it is important to estimate the remainder of the series in
order to bound the amount by which \ref{bootstrapequation} might be violated in
a given approximation. This was done by Rychkov et al,
who found the following error estimate holds (for $\Delta_*\gg 1$):
\begin{equation}
\Big|\sum_{(\Delta\geq \Delta_*),l}p_{\mathcal{O}} \, g_{\Delta,l}(z,\bar z)\Big| \, \leq \, \mathcal{R}(z,\bar z)\,,
\label{finalest4}
\end{equation}
where
\begin{equation}
\mathcal{R}(z,\bar z) \,\equiv \, \frac{(-\log|\rho(z)|)^{-4\Delta_\phi+1} \, 2^{4\Delta_\phi+1}}{\Gamma(4\Delta_\phi)} \frac{\Gamma(4\Delta_\phi, - \Delta_* \log |\rho(z)|)}{{(1-|\rho(z)|^2)}}\,.
\label{finalest5a}
\end{equation}

@@

The straightforward path is to define a cost function
\[
  C\sim | \mathcal{M}\cdot \vec{\rho} - \vec{\sigma} |
\]
and minimize it stochastically\footnote{Should I add a section on Metropolis MC
here?}.
This naive approach proved to be untractable in the early stages of the
research, mostly because it became exponentially time consuming to look for more
operators. Thus, we decided to split the optimization in two steps. One that
finds an approximate spectrum without worrying about the OPE coefficients, and
then solving for the complete conformal data with a more refined Monte Carlo.
The first step was motivated by Gliozzi's determinant bootstrap, which we review
in the following section.

@@

\section{LogDet implementation}
\subsection{naive approach}
Given a putative spectrum, the determinant of \ref{matrix-op}
should get closer to zero as we approach a physically sensible set of operator
dimensions. Since this quantity could range over many orders of magnitude, we
decided to square it and then take the $\log$ of it. This proved to be
numerically troubling due to the behaviour of the logarithm around $0$  and the landscape was thus full of spurious minima out of
which the Metropolis algorithm could not get out.

\subsection{Cross smearing}
In order to smooth this behaviour out, we implemented an empirical
renormalization that consists in computing the value of the determinant at
adjacent points to the main one. Thus
\[
\begin{aligned}
  S({\Delta_i,l_i})& = \log\Det^2\mathcal{M}(\{\Delta_i,l_i\})\\
  \to
  S_{\mathrm{cross}}(\{\Delta_i,l_i\})  & =   \log\Det^2\mathcal{M}(\{\Delta_i,l_i\}) + \sum_i\left(
  \log\Det^2\mathcal{M}(\cdots,\{\Delta_i+\delta,l_i\},\cdots)  +
  \log\Det^2\mathcal{M}(\cdots,\{\Delta_i-\delta,l_i\},\cdots)  \right)
\end{aligned}
\]

This action was much more efficient at finding the correct dimensions for the
free theory, and numerical exploration showed that the basin of attraction for
each given operator's dimension was very sharp, as long as the other ones were
kept at the exact value. Small perturbations destroyed this basin rather
quickly, and thus it was hard to get the MC routine to solve for a large number
of operators  ($\sim 7$) at the same time.

@@

The next breakthrough in the routine came when we decided to start with few
($\sim 4$) operators, find the operator dimensions which minimized the 
action and then add the next operators one by one. To avoid spoiling
the previous solution, only the new operator was varied for a given number of
steps before we allowed the whole spectrum to be optimized. This gave a very
efficient routine with a large radius of convergence.

@@

\section{$\chi^2$ refinement}
\subsection{cWLS as a way to determine the OPE coefficients}
Since the dependence of \ref{cross-check} on $\vec{\rho}$ is linear, for
$N_z>N_{Op}$ it is possible to find the set of OPE coefficients that best
satisfy crossing via the Weighted Least Squares method\footnote{Should we add a
reference here, other than Wikipedia?}. 
For a generic spectrum
obtained via the method described in section \ref{rec-check}, this naive set of OPE
coefficients will have some negative components, which stems from the
fact that our spectrum is only approximate. Since the error estimate
\ref{finalestimate5a}
depends
crucially on the positivity of the OPE coefficients, we need to find the
solution that minimizes \ref{cross-check} and has all entries non-negative. This
is attained by setting the first negative OPE coefficient (from back to forth,
if $\Delta_i<\Delta_{i+1}$) and then running a WLS routine with the remaining
entries. If this solution still has negative entries, the proccedure is repeated
until an acceptable $\vec{\rho}$ is obtained.
